The robot who pretend-played with me

### 2017-03-27 paper.tex
Re-structuring the notes on the model of social awareness in a paper on freeplay

\title{From Social Attention to Social Modelling in Robotics:\\ The Robotic Social Awareness Model}

\author{Séverin Lemaignan}

@@

\begin{abstract}
Abstract goes here
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{What do we aim for?}

@@

A principled yet workable approach to social cognition for artificial intelligence

\subsection{Problem Statement}

@@

\subsection{From Social Attention to Social Modelling}

The hypothesis that we hereafter develop and turn into a cognitive model is the
following: mental representations are snapshots of awareness, awareness being
itself a label for the memory-mediated process of attention.

\subsection{Surface Alignment vs Deep Grounding}

\subsection{Theory of Mind}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Foundations of the Socio-cognitive Awareness for Robots}

\subsection{Connections vs. representations}
\label{connection-representation}

@@

\subsection{The Attention Schema Theory}

@@

\subsection{Associative Memory Networks}

In particular, used for behavioural alignment

\subsection{Semantic Networks}

Ontologies and other Knowledge Representation and Reasoning systems.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{The Robotic Social Awareness Model}

\subsection{Overview}

@@

\subsection{Attention}

@@

\begin{figure}
    \centering
        \resizebox{0.7\linewidth}{!}{%
        \begin{tikzpicture}[
                >=latex,
                every edge/.style={draw, very thick},
                every node/.style={font=\bf,align=center},
                input/.style={draw,circle, inner sep=0pt,minimum size=0.6cm}
            ]

@@

\subsubsection{Biasing Mechanisms}

@@

\subsection{Social Attention}

@@

\subsection{What is the contribution, the novelty?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{What Does the Model Predict?}

\subsection{Behavioural Alignment}
\subsection{Verbalization of the (Social) Attentional Process}
\subsection{False-beliefs}

\subsection{Symbol Grounding}

\subsection{Recursive Awareness}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Functional Pre-Requisites for Social Robots}

\subsection{Social Assessment}
\subsubsection{Gaze Direction}
\subsubsection{Facial and Body Expressions}
\subsubsection{Gestures}

\subsection{Situation Assessment}

Including saliency information

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Evaluation of the Model}

\subsection{Does it perform as well as Symbolic Architectures?}
\subsection{Comparison with Symbolic Situation Assessment}

>>>>

\graphicspath{{figs/}}

\title{The robot who pretend-played with me}

\author{Séverin Lemaignan, Tony Belpaeme}

@@

Alternative titles:

\begin{itemize}
    \item Can I have real fun playing with my robot?
    \item A Window Onto Your Mind: Playful Robots Leveraging Attention to
        Predict Actions
    \item Playing Like your Friend: Dealing with the High Dimensionality of
        Real-World Social Dynamic

\end{itemize}

\begin{abstract}

    This article presents how attention tracking enables a robot to first automatically segment
    a social interaction into meaningful situations (\emph{situation awareness}), and
    second, to predict the participants' actions.

    We study this effect in a high-dimensional social interaction situation based on
    free dramatic play: pairs of children (4-7 years old) or robot and child are
    invited to freely interact with characters and items on an interactive
    table, telling and acting stories with no explicit preset goal.

    Although the actions performed by the children are not known beforehand, our
    approach allows the robot to segment on the fly the interaction flow into
    episodes whose main focus can be predicted and used to generate pro-social
    behaviours.

\end{abstract}

\section{Introduction}

\subsection{Mutual Modelling: Theoretical Background}

@@

A principled yet practical approach to social cognition for artificial intelligence

\subsubsection{Pretense}

A second direction for the 'meta-cognitive foundations of mind modelling' is
Leslie's paper on the role of pretense~\cite{leslie1987pretense}.

\subsubsection{Surface Alignment vs Deep Grounding}

analysis of behavioural alignment between partners (via
metrics like the recently proposed \emph{Individual Motor
Signature}~\cite{slowinski2016dynamic})

\subsubsection{Theory of Mind}

\subsubsection{Connections vs. representations}
\label{connection-representation}

\subsection{Attention in Robotics}

\section{Methodology: Social Interaction during Free Play}

The study of the emergence of social behaviour is often intrinsically limited by
the nature of the experimental tasks. We introduce a new
experimental paradigm based on free play interactions. Where traditional
socio-cognitive tasks are either toy scenarios (\ie they do not adequately
mirror real-world situations) or simple, constrained tasks that do not reflect
the complexity and dynamics of real-world interactions, free play social
interactions exhibit a rich set of natural cognitive and social dynamics that
are desirable for this research.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{setup.png}
    \caption{The free play social interactions sandbox}
    \label{fig|setup}
\end{figure}

To mitigate the technical challenges presented by free play interactions (like
high dimensionality and uncertainties), we have developped a  \emph{free
play sandbox}~\ref{fig|setup}: open-ended and non-directive activities, yet sufficiently
defined to be reproducible; focus on abstract socio-cognitive facets (robot
perception and manipulation are simplified); well suited for qualitative and
quantitative analysis using metrics like Słowinski’s Individual Motor Signature
(for behavioural alignment), Ballard’s coding of children’s free-play
interactions, with-me-ness (for assessment of co-engagment).

\subsection{Free play task}

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{sandbox}
    \caption{Example of a game situation: the child and the robot `build'
    together a zoo. The animals and the blocks can be dragged over the whole
    play area, while the background picture is static and purely decorative.}
    \label{fig|sandbox}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.7\linewidth]{sandbox-release-cheetah}
    \caption{Possible play situation: the child decides to release the cheetah
    and creates a story where the cheetah scares away the visitors.}
    \label{}
\end{figure}

\subsection{Data Analysis}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{analysis}
    \caption{Screenshot of the interaction analysis tool: faces of the two
    children are recorded, as well as any interaction occuring on the
    interactive table.}
    \label{fig|analysis}
\end{figure}

\section{Discussion: Towards a Model of Socio-Cognitive Awareness for Robots}

\subsection{The classical approach}

@@

[What is not adequately solved by current techniques?]

[give concrete examples of social situations that are not easily achieveable
with current (symbolic or not) approaches]

[this examples should be turned into predictions for what our proposal should be
able to achieve]

@@

\subsection{From Social Attention to Social Modelling: the Attention Schema Theory}

The hypothesis that we hereafter develop and turn into a cognitive model is the
following: \textbf{mental representations are snapshots of awareness, awareness being
itself a label for the memory-mediated process of attention}.

This extends to social cognition: \textbf{modelling others' mental representations is
taking snapshots of their current state of awareness}. As we do not have direct
access to others' process of attention, it has to be mediated. We suggest that
\textbf{modelling other's state of awareness is mediated by one's own
attentional system, through joint attention mechanisms}.

@@

@@

\paragraph{Attention}

@@

\paragraph{Memory-mediated Attention}

\emph{Attention} is modelled as the set of activated units in an associative
memory network.

Interplay between multi-modal social cues on one hand, and priming of previously activated physical
entities (objects, agents) and

Attention

\begin{figure}[h!]
    \centering
        \resizebox{0.7\linewidth}{!}{%
        \begin{tikzpicture}[
                >=latex,
                every edge/.style={draw, very thick},
                every node/.style={font=\bf,align=center},
                input/.style={draw,circle, inner sep=0pt,minimum size=0.6cm}
            ]

@@

\paragraph{Biasing Mechanisms}

@@

\paragraph{Social Attention}

@@

\subsection{What Does the Model Predict?}

\paragraph{Protodeclarative pointing}

Protoimperative vs protodeclarative pointing~\cite{baron1986perceptual}

\paragraph{Behavioural Alignment}
\paragraph{Verbalization of the (Social) Attentional Process}
\paragraph{False-beliefs}

Picture ordering protocol, with \emph{mechanical}, \emph{behavioural} and
\emph{intentional} sequences~\cite{baron1986mechanical}

\paragraph{Symbol Grounding}

\paragraph{Recursive Awareness}

@@

\paragraph{Adapted behaviour to the interacting agent}

A robot facing a baby, a 3 years old, a 5 years old, a 13 years old or an adult
should behave differently.

\paragraph{Parten's stages of play}

\section{Conclusion}
